\sys: Light-touch Data-Parallel Shell Processing

### 2020-07-16 pldi/draft.tex
Prepare arxiv version

@@

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.
%% Author with single affiliation.
\author{
  Anonymous Author(s)\\
  \normalsize{Additional Anonymized Material: \href{https://git.io/JfwSq}{https://git.io/JfwSq}.}
}
%% \authornote{with author1 note}          %% \authornote is optional;
%%                                         %% can be repeated if necessary
%% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
%% \affiliation{
%%   \position{Position1}
%%   \department{Department1}              %% \department is recommended
%%   \institution{Institution1}            %% \institution is required
%%   \streetaddress{Street1 Address1}
%%   \city{City1}
%%   \state{State1}
%%   \postcode{Post-Code1}
%%   \country{Country1}                    %% \country is recommended
%% }
%% \email{first1.last1@inst1.edu}          %% \email is recommended

@@

The \unix shell is an environment---often interactive---for composing scripts written in a plethora of programming languages.
This language-agnosticism, coupled with \unix's toolbox philosophy~\cite{mcilroy1978unix}, makes the shell the primary choice for specifying succinct and simple pipelines for data processing, system orchestration, and other automation tasks.
% Pipe-centric shell commands are parallel (and to some extent, functional) programs that can natively enjoy significant parallelism on a multicore machine.
Unfortunately, parallelizing such pipelines requires significant effort shared between two different programmer groups. % it's more like roles personas or "wearing hats"

@@

\sys defines aggregators for many \pur POSIX and GNU commands, which operate as both \sys's standard library and an exemplar for community attempts to tackle other commands.

@@

The vast majority requires no effort other than invoking \sys;
  the annotation for the remaining \todo{6} commands ($<3\%$) amounts to a single record.
% 
In terms of correctness, \sys's results over multi-GB inputs are identical to the sequential for all benchmarks.
Scripts feature ample opportunities for breaking semantics, which \sys avoids.
%% In one case, \sys avoids transforming a cycle (\ie a fragment whose input and output identifiers are the same).
%in another, it preserves the semantics that another parallelization command breaks.

@@

There are a few possible ways one might attempt to use GNU \ttt{parallel} on this program.
They could use it on the 8th stage, assuming they know it is a bottleneck, bringing execution down to 304.4s ($1.8\times$ speedup).
Alternatively, they could (incorrectly) sprinkle \ttt{parallel} across the entire program---a strategy simplified by \ttt{parallel}'s \ttt{stdin} redirection feature.
This would lead to 3.2$\times$ performance improvements but severely incorrect results with respect to the sequential execution---with 92\% of the output showing a difference between sequential and parallel execution.
\sys's conservative program transformations will not attempt parallelizing program sections that contain commands with unclear parallelizability properties.
	% or commands whose inputs depend on earlier outputs.
% GNU \ttt{parallel} is oblivious to command semantics, risking breaking program semantics as in this case.

@@

%% \begin{acks}
  % Dumping people so that we don't forget
  % 
  %% This material is based upon work supported by the
  %% \grantsponsor{GS100000001}{National Science
  %%   Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  %% No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  %% No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  %% conclusions or recommendations expressed in this material are those
  %% of the author and do not necessarily reflect the views of the
  %% National Science Foundation.
%% \end{acks}

%% Bibliography
\bibliography{./bib}

>>>>

\newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}

@@

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.
%% Author with single affiliation.
\author{
% Anonymous Author(s)\\
% \normalsize{Additional Anonymized Material: \href{https://git.io/JfwSq}{https://git.io/JfwSq}.}
  {\rm \normalsize Nikos Vasilakis$^{*}$\qquad Konstantinos Kallas$^{*\dagger}$\qquad  Konstantinos Mamouras$^{\circ}$ }\\
  {\rm \normalsize Achilleas Benetopoulos$^{\diamond}$ \qquad Lazar Cvetkovich$^{\ddagger}$ }\\
  {\small MIT, CSAIL \quad $^{\dagger}$University of Pennsylvania \quad $^{\circ}$Rice University \quad $^{\diamond}$Unaffiliated \quad $^{\ddagger}$University of Belgrade}
}
% \authornote{$^{*}$Equal contribution.}     %% \authornote is optional;
%%                                         %% can be repeated if necessary
%% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
%% \affiliation{
%%   \position{Position1}
%%   \department{Department1}              %% \department is recommended
%%   \institution{Institution1}            %% \institution is required
%%   \streetaddress{Street1 Address1}
%%   \city{City1}
%%   \state{State1}
%%   \postcode{Post-Code1}
%%   \country{Country1}                    %% \country is recommended
%% }
%% \email{first1.last1@inst1.edu}          %% \email is recommended

@@

\blfootnote{$^{*}$Equal contribution.}
The \unix shell is an environment---often interactive---for composing scripts written in a plethora of programming languages.
This language-agnosticism, coupled with \unix's toolbox philosophy~\cite{mcilroy1978unix}, makes the shell the primary choice for specifying succinct and simple pipelines for data processing, system orchestration, and other automation tasks.
% Pipe-centric shell commands are parallel (and to some extent, functional) programs that can natively enjoy significant parallelism on a multicore machine.
Unfortunately, parallelizing such pipelines requires significant effort shared between two different programmer groups. % it's more like roles personas or "wearing hats"

@@

\sys defines aggregators for many POSIX and GNU commands in \pur, dubbing as both its standard library and an exemplar for community efforts tackling other commands.

@@

The vast majority requires no effort other than invoking \sys;
  the annotation for the remaining 6 commands ($<3\%$) amounts to a single record.
% 
In terms of correctness, \sys's results over multi-GB inputs are identical to the sequential for all benchmarks.
Scripts feature ample opportunities for breaking semantics, which \sys avoids.
%% In one case, \sys avoids transforming a cycle (\ie a fragment whose input and output identifiers are the same).
%in another, it preserves the semantics that another parallelization command breaks.

@@

There are a few possible ways one might attempt to use GNU \ttt{parallel} on this program.
They could use it on the 8th stage, assuming they know it is a bottleneck, bringing execution down to 304.4s ($1.8\times$ speedup).
Alternatively, they could (incorrectly) sprinkle \ttt{parallel} across the entire program---a strategy simplified by \ttt{parallel}'s \ttt{stdin} redirection feature.
This would lead to 3.2$\times$ performance improvements but severely incorrect results with respect to the sequential execution---with 92\% of the output showing a difference between sequential and parallel execution.
\sys's conservative program transformations will not operate on fragments with unclear parallelizability properties.
	% or commands whose inputs depend on earlier outputs.
% GNU \ttt{parallel} is oblivious to command semantics, risking breaking program semantics as in this case.

@@

\subsection*{Acknowledgements}
% Dumping people so that we don't forget
We want to thank 
  Andr\'e DeHon,
  Ben Karel,
  Jean-S\'ebastien L\'egar\'e, and 
  Nick Roessler.
This material is based upon work supported by DARPA contract no. HR00112020013 (ALRSCC) and NSF award CCF 1763514.
%% \grantsponsor{GS100000001}{National Science
%%   Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
%% No.~\grantnum{GS100000001}{nnnnnnn} and Grant
%% No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
%% conclusions or recommendations expressed in this material are those
%% of the author and do not necessarily reflect the views of the
%% National Science Foundation.

%% Bibliography
{\small
\bibliography{./bib}
}
